Forecasting the 2024 U.S. Presidential Election: A Poll-Based Approach

### 2024-11-03 paper/paper.qmd
Updated introduction and abstract

---
title: "Forecasting the 2024 U.S. Presidential Election: A Poll-Based Approach"
subtitle: "Kamala Harris Poised for Victory"
author: 
  - Xinxiang Gao
  - Ariel Xing
  - John Zhang^[The authors are listed in alphabetical order by last name.]
thanks: "Code and data are available at: [https://github.com/xgao28/election_forecast](https://github.com/xgao28/election_forecast)."
date: today
date-format: long
abstract: "The upcoming U.S. Presidential Election on November 5th features a closely contested race between Vice President Kamala Harris and former President Donald Trump. This paper uses a poll-of-polls approach, combined with a dynamic linear regression model that updates daily with the latest data, to forecast the popular vote and outcomes in seven battleground states: Arizona, Georgia, Nevada, North Carolina, Michigan, Wisconsin, and Pennsylvania. Our analysis currently projects that Vice President Harris will win the popular vote, 48.19% to 46.01%, and secure the electoral college with 302 votes to Trump’s 236, by winning five of these seven swing states: Nevada, North Carolina, Michigan, Wisconsin, and Pennsylvania. By aggregating multiple polls rather than relying on a single poll, our model provides a more robust prediction, supporting the likelihood of a Harris victory."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

We apply a dynamic linear regression model named “ Baseline Model” to provide a national overview and another dynamic linear regression model named “Primary Model” to assess state-level dynamics, especially in key battlegrounds. The Baseline Model aggregates national polling data to estimate popular vote share, while the Primary Model offers a state-specific analysis aligned with the electoral college structure. This approach addresses a gap in existing models that often overlook the unique impact of swing states on electoral outcomes.
Our model projects that Vice President Harris will secure 48.19% of the popular vote, compared to 46.01% for Trump. However, as U.S. elections are decided by the electoral college rather than the popular vote, a candidate can win the presidency without a majority of popular support. There are seven swing states, Arizona, Georgia, Nevada, North Carolina, Michigan, Wisconsin, and Pennsylvania, holding a collective 93 electoral votes, are determining factors to the election’s outcome. Excluding swing states, Harris would receive 226 electoral votes, while Trump would receive 219. According to our Primary Model, Harris is expected to capture five of the seven swing states: Nevada, North Carolina, Michigan, Wisconsin, and Pennsylvania, yielding a total of 302 electoral votes and thus securing the presidency. Trump would win two swing states, Arizona and Georgia, ending with 236 electoral votes. Although Harris's lead in North Carolina is narrow (by just 0.036%), which may switch to voting for Trump in the future, but 19 votes in this state in the face of overwhelming vote margins won't have much of an impact on the outcome.

>>>>

---
title: "Forecasting the 2024 U.S. Presidential Election: A Poll-Based Approach"
subtitle: "Kamala Harris Poised for Victory"
author: 
  - Xinxiang Gao
  - Ariel Xing
  - John Zhang^[The authors are listed in alphabetical order by last name.]
thanks: "Code and data are available at: [https://github.com/xgao28/election_forecast](https://github.com/xgao28/election_forecast)."
date: today
date-format: long
abstract: "The upcoming U.S. Presidential Election on November 5th features a closely contested race between Vice President Kamala Harris and former President Donald Trump. This paper uses a poll-of-polls approach, combined with a dynamic linear regression model that updates daily with the latest data, to forecast the popular vote and outcomes in seven battleground states: Arizona, Georgia, Nevada, North Carolina, Michigan, Wisconsin, and Pennsylvania. Our analysis currently projects that Vice President Harris will win the popular vote, 48.19% to 46.01%, and secure the electoral college with 292 votes to Trump’s 246, by winning five of these seven swing states: Nevada, North Carolina, Michigan, Wisconsin, and Pennsylvania. By aggregating multiple polls rather than relying on a single poll, our model provides a more robust prediction, supporting the likelihood of a Harris victory."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

We apply a dynamic linear regression model named “ Baseline Model” to provide a national overview and another dynamic linear regression model named “Primary Model” to assess state-level dynamics, especially in key battlegrounds. The Baseline Model aggregates national polling data to estimate popular vote share, while the Primary Model offers a state-specific analysis aligned with the electoral college structure. This approach addresses a gap in existing models that often overlook the unique impact of swing states on electoral outcomes.
Our model projects that Vice President Harris will secure 48.19% of the popular vote, compared to 46.01% for Trump. However, as U.S. elections are decided by the electoral college rather than the popular vote, a candidate can win the presidency without a majority of popular support. There are seven swing states, Arizona, Georgia, Nevada, North Carolina, Michigan, Wisconsin, and Pennsylvania, holding a collective 93 electoral votes, are determining factors to the election’s outcome. Excluding swing states, Harris would receive 226 electoral votes, while Trump would receive 219. According to our Primary Model, Harris is expected to capture five of the seven swing states: Nevada, North Carolina, Michigan, Wisconsin, and Pennsylvania, yielding a total of 292 electoral votes and thus securing the presidency. Trump would win two swing states, Arizona and Georgia, ending with 246 electoral votes. Although Harris's lead in North Carolina is narrow (by just 0.036%), which may switch to voting for Trump in the future, but 19 votes in this state in the face of overwhelming vote margins won't have much of an impact on the outcome.
